---
title: "What Makes the Best Tennis Player?"
author: "Steven Herrera and Ethan Shen"
date: "11/09/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, message=F,include=F}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')
```

```{r load-packages, message=FALSE}
library(tidyverse)
```

# Introduction

# Data 

We used the following data set: `atp_matches_2017`, created by JeffSackMann on GitHub.  
Here is a link to the dataset: [https://github.com/JeffSackmann/tennis_atp/blob/master/atp_matches_2017.csv](https://github.com/JeffSackmann/tennis_atp/blob/master/atp_matches_2017.csv).  
  
The dataset that we will be using for our analysis is not the same as the one we initially received, as multiple dataset manipulation techniques were used in Excel and in R. Look at **Additional Work**.

Here is a glimpse of our dataset:
```{r}

```




# Analysis 

1) Description of the response variable and its variable type

2) Description of the explanatory variables and the population coefficients you wish to understand using statistical inference

possible explanatory variables: winner hand, winner age, winner rank, winner rank points

3) Hypothesis/hypotheses regarding your question of interest

4) Proposed methods to use for the analysis 

# Reference 



# Additional Work 

1) In order to obtain the dataset in our RMD, we will first need the data to fit nicely into a CSV file. In excel, we used the `Text to Columns` function, and inserted the file here:

```{r message=FALSE}
atp <- read_csv("files/atp.csv")
```

2) There are a lot of things interesting about this dataset. The first thing that we did was filter for events that are normal tournament-style matches (not ATP Final events or Davis Cup, which both occur past November 13, 2017, coded as `20171113` in the dataset under the `tourney_name` variable).

```{r}
atp1 <- atp %>%
  filter(tourney_date < 20171113)
```

3) Because we are interested in winners, we will only look at the observations where the match was a **final**.

```{r}
winners <- atp1 %>%
  filter(round == "F")
```

4) Because we are looking in the time period from 2010-2017, we will do the same thing for the rest of the years:

```{r message=FALSE}
atp2016 <- read_csv("files/atp2016.csv")
atp2015 <- read_csv("files/atp2015.csv")
atp2014 <- read_csv("files/atp2014.csv")
atp2013 <- read_csv("files/atp2013.csv")
atp2012 <- read_csv("files/atp2012.csv")
atp2011 <- read_csv("files/atp2011.csv")
atp2010 <- read_csv("files/atp2010.csv")

winners2016 <- atp2016 %>%
  filter(tourney_date < 20161114) %>%
  filter(round == "F")

winners2015 <- atp2015 %>%
  filter(tourney_date < 20151115) %>%
  filter(round == "F")

winners2014 <- atp2014 %>%
  filter(tourney_date < 20141109) %>%
  filter(round == "F")

winners2013 <- atp2013 %>%
  filter(tourney_date < 20131104) %>%
  filter(round == "F")

winners2012 <- atp2012 %>%
  filter(tourney_date < 20121105) %>%
  filter(round == "F")

winners2011 <- atp2011 %>%
  filter(tourney_date < 20111114) %>%
  filter(round == "F")

winners2010 <- atp2010 %>%
  filter(tourney_date < 20101121) %>%
  filter(round == "F")
```


