---
title: "Assumptions"
author: "Steven Herrera and Ethan Shen"
date: "11/09/2018"
output:
  pdf_document: default
  html_document: default
---

```{r load-packages, message=FALSE, include=FALSE}
library(tidyverse)
library(olsrr)
library(cowplot)
library(car)
library(broom)
library(knitr)
library(arm)
library(tidyr)
library(pROC)
library(arm)
library(rlm)
```

```{r message=FALSE, include=FALSE}
atp <- read_csv("files/atp.csv")
atp2016 <- read_csv("files/atp2016.csv")
atp2015 <- read_csv("files/atp2015.csv")
atp2014 <- read_csv("files/atp2014.csv")
atp2013 <- read_csv("files/atp2013.csv")
atp2012 <- read_csv("files/atp2012.csv")
atp2011 <- read_csv("files/atp2011.csv")
atp2010 <- read_csv("files/atp2010.csv")

atp1 <- atp %>%
  filter(tourney_date < 20171113)

winners2016 <- atp2016 %>%
  filter(tourney_date < 20161114)

winners2015 <- atp2015 %>%
  filter(tourney_date < 20151115) 

winners2014 <- atp2014 %>%
  filter(tourney_date < 20141109) 

winners2013 <- atp2013 %>%
  filter(tourney_date < 20131104) 

winners2012 <- atp2012 %>%
  filter(tourney_date < 20121105) 

winners2011 <- atp2011 %>%
  filter(tourney_date < 20111114) 

winners2010 <- atp2010 %>%
  filter(tourney_date < 20101121) 

winners <- rbind(atp1, winners2016, winners2015, winners2014, winners2013, 
                 winners2012, winners2011, winners2010)

winners_set <- winners %>%
  filter(best_of == 3,
         !is.na(w_ace), 
         !is.na(w_df), 
         !is.na(w_svpt), 
         !is.na(w_1stIn), 
         !is.na(w_1stWon), 
         !is.na(w_2ndWon),
         !is.na(w_SvGms), 
         !is.na(w_bpSaved),
         !is.na(w_bpFaced),
         surface!="None")

losers <- winners_set %>%
  mutate(seed = loser_seed,
         name = loser_name,
         hand = loser_hand,
         ht = loser_ht,
         age = loser_age,
         rank = loser_rank,
         rankpoints = loser_rank_points,
         ace = l_ace,
         df = l_df,
         svpt = l_svpt,
         firsstIn = l_1stIn,
         firsttWon = l_1stWon,
         secondndWon = l_2ndWon,
         SvGms = l_SvGms,
         bpSaved = l_bpSaved,
         bpFaced = l_bpFaced,
         minutes = minutes,
         status = 0
         )  
  

winners <- winners_set %>%
  mutate(seed = winner_seed,
         name = winner_name,
         hand = winner_hand,
         ht = winner_ht,
         age = winner_age,
         rank = winner_rank,
         rankpoints = winner_rank_points,
         ace = w_ace,
         df = w_df,
         svpt = w_svpt,
         firsstIn = w_1stIn,
         firsttWon = w_1stWon,
         secondndWon = w_2ndWon,
         SvGms = w_SvGms,
         bpSaved = w_bpSaved,
         bpFaced = w_bpFaced,
         minutes = minutes,
         status = 1
         ) 
tennis <- rbind(winners,losers)

tennis <- tennis %>%
  filter(!is.na(seed),
         !is.na(name),
         !is.na(hand),
         !is.na(ht),
         !is.na(age),
         !is.na(rank),
         !is.na(rankpoints),
         !is.na(ace),
         !is.na(df),
         !is.na(svpt),
         !is.na(firsstIn),
         !is.na(firsttWon),
         !is.na(secondndWon),
         !is.na(SvGms),
         !is.na(bpSaved),
         !is.na(bpFaced),
         !is.na(minutes),
         !is.na(status))
```

```{r include=FALSE, message=FALSE}
glimpse(tennis)
```

```{r include=FALSE, message=FALSE}
set.seed(1234)
ten <- tennis %>% sample_n(1000)
```

```{r include=FALSE, message=FALSE}
glimpse(ten)
```

# Logistic Regression Assumptions

After looking at exploratory data analysis, found on our Project RMD file, we have concluded with the following model that we will use as our final model:

```{r}
newten <- ten
final <- glm(status ~ minutes + ht + rankpoints + ace + df, family = binomial, 
             data = newten)
kable(tidy(final), format = "markdown", digits = 6)
```

# Model Assessment

## Binned Plots with Residuals vs Predicted

We will further our investigation of whether our new model follows the key model assessment characteristics:

- Good binned residual vs. predicted plot

- Good binned residual vs. numerical explanatory plots

- Large area under the ROC curve

```{r}
newten <- newten %>% mutate(Residuals = residuals.glm(final,type="response"),
                          Predicted = predict.glm(final,type="response"))

binnedplot(newten$Predicted, newten$Residuals,xlab="Predicted Probabilities",
           ylab="Residuals",main="Binned Residuals vs. Predicted Probabilities")
```

```{r,fig.height=4,fig.width=4,echo=F}
binnedplot(newten$minutes, newten$Residuals,xlab="Minutes",
           ylab="Residuals",main="Binned Residuals vs. Minutes")

binnedplot(newten$ht, newten$Residuals,xlab="Height",
           ylab="Residuals",main="Binned Residuals vs. Height")

binnedplot(newten$ace, newten$Residuals,xlab="Aces",
           ylab="Residuals",main="Binned Residuals vs. Aces")

binnedplot(newten$df, newten$Residuals,xlab="Double Faults",
           ylab="Residuals",main="Binned Residuals vs. Double Faults")

binnedplot(newten$rankpoints, newten$Residuals,xlab="Rankpoints",
           ylab="Residuals",main="Binned Residuals vs. Rankpoints")
```

```{r}
ROC.newten <- roc(newten$status,newten$Predicted,plot=T)
```

```{r}
ROC.newten$auc
```

# Influential Points 

```{r}
newten <- newten %>%
  mutate(leverage = hatvalues(final), 
         cooks = cooks.distance(final),
         stand.resid = rstandard(final), 
         obs.num = row_number())
```

## Leverage and Cook's Distance 
```{r fig.height=2}
ggplot(data=newten, aes(x=obs.num,y=leverage)) + 
  geom_point(alpha=0.5) + 
  geom_hline(yintercept=0.1,color="red")+
  labs(x="Observation Number",y="Leverage",title="Leverage")
```

```{r fig.height=2}
ggplot(data=newten, aes(x=obs.num,y=cooks)) + 
  geom_point() + 
  geom_hline(yintercept=1,color="red")+
  labs(x="Observation Number",y="Cook's Distance",title="Cook's Distance")
```

## VIF and Multicollinearity 

```{r}
tidy(vif(final))
```

```{r}
pairs(status ~ minutes + ht + rankpoints + ace + df, data = newten)
```

# Conclusion

With VIF values less than 10 and a scatterplot matrix that does not show obvious linear relationships between the explanatory variables, observations that are under the leverage and Cook's distance threshold, and binned residual plots that satisfy the assumptions, we have cleared model assessment and assumptions for the following final model:

```{r}
kable(tidy(final), format = "markdown", digits = 6)
```




